---
title: "Notes5"
author: "nisha pardeshi"
date: "1 October 2018"
output: html_document
---

Interval Estimation for Population Mean
```{r}
# Xbar = 102, n=50, sigma=10, what is 95% CI of mu
Xbar = 102
n=50
sigma=10
alpha=1-0.95

#Zalpha/2 
CV = qnorm(alpha/2)
paste("CV", CV)

SE = sigma/sqrt(n)
paste("SE", SE)

LL = Xbar - CV * SE
UL = Xbar + CV * SE
paste("LL", UL)
paste("UL", LL)

```
```{r}
# Xbar = 102, n=50, sigma=10, what is 99% CI of mu
Xbar = 102
n=50
sigma=10
alpha=1-0.99

#Zalpha/2 
CV = qnorm(alpha/2)
paste("CV", CV)

SE = sigma/sqrt(n)
paste("SE", SE)

LL = Xbar - CV * SE
UL = Xbar + CV * SE
paste("LL", UL)
paste("UL", LL)
```
```{r}
# Xbar = 102, n=50, sigma=10, what is 100% CI of mu
Xbar = 102
n=50
sigma=10
alpha=1-1

#Zalpha/2 
CV = qnorm(alpha/2)
paste("CV", CV)

SE = sigma/sqrt(n)
paste("SE", SE)

LL = Xbar - CV * SE
UL = Xbar + CV * SE
paste("LL", UL)
paste("UL", LL)
```

Interval Estimation for Unknown Sigma
```{r}
#Xbar = 81, s=30, n=16 
# Assume population normally distributed
# 95% of CI of mu
Xbar = 81
s = 30
n = 16
alpha = 1 - 0.95
# SInce sigma is unknown we go for t distribution
CV = qt(alpha/2, n-1)
paste("CV", CV)

SE = s/sqrt(n)
paste("SE", SE)

LL = Xbar - CV * SE
UL = Xbar + CV * SE
paste("LL", UL)
paste("UL", LL)
```
```{r}
# Assume population normally distributed
a <- c(1,4,4,4,5,5,5,8)
b <- 1:8

# Construct 95% CI of a
Xbar = mean(a)
s = sd(a)
n = length(a)
paste("Xbar", Xbar)
alpha = 1 - 0.95
# SInce sigma is unknown we go for t distribution
CV = qt(alpha/2, n-1)
paste("CV", CV)

SE = s/sqrt(n)
paste("SE", SE)

LL = Xbar - CV * SE
UL = Xbar + CV * SE
paste("LL", UL)
paste("UL", LL)
```
```{r}
# Construct 95% CI of b
Xbar = mean(b)
s = sd(b)
n = length(b)
alpha = 1 - 0.95
paste("Xbar", Xbar)
# SInce sigma is unknown we go for t distribution
CV = qt(alpha/2, n-1)
paste("CV", CV)

SE = s/sqrt(n)
paste("SE", SE)

LL = Xbar - CV * SE
UL = Xbar + CV * SE
paste("LL", UL)
paste("UL", LL)

# why a and b produce different CI even if mean and range are same
# Because SE differs in both case
```
```{r}
mpg <- c(20,34,22,26,32,23,20,28,33,32,35,31,29,24)
# Construct 99% CI, assuming normal distribution
Xbar = mean(mpg)
s = sd(mpg)
n = length(mpg)
alpha = 1 - 0.99
paste("Xbar", Xbar)
# SInce sigma is unknown we go for t distribution
CV = qt(alpha/2, n-1)
paste("CV", CV)

SE = s/sqrt(n)
paste("SE", SE)

LL = Xbar - CV * SE
UL = Xbar + CV * SE
paste("LL", UL)
paste("UL", LL)
```

Interval Estimation for population proportion
```{r}
# Of 100 exec surveyed, 93 stated that sal of top mangmt should be based on performance.
phat = 93/100
n = 100
phat * n > 10
(1-phat)*n > 10
n > max(10/phat, 10/(1-phat))
# n is not large enough CI is not reliable

# Construct 95% CI


# Assume phat is normally distributed
alpha = 1 - 0.95
CV = qnorm(alpha/2)
paste("CV", CV)
SE = sqrt(phat*(1 - phat)/n)
paste("SE", SE)

LL = phat - CV * SE
UL = phat + CV * SE
paste("LL", UL)
paste("UL", LL)

```

```{r}
#Estimate proportion of househole to purchase tel line,
#Data for additional household is random sample of 500 house.
# Results indicate 175 would purchase.
phat = 175/500
n = 500
phat*n > 10
(1-phat)*n > 10
n > max(10/phat, 10/(1-phat))
# n is large enough CI is reliable
# Construct 99% CI the would purchase
alpha = 1-0.99
CV = qnorm(alpha/2)
paste("CV", CV)
SE = sqrt(phat*(1-phat)/n)
paste("SE", SE)
LL = phat - CV * SE
UL = phat + CV * SE
paste("LL", UL)
paste("UL", LL)

# Between 30% and 40% chances are they would buy
```
Determining Sample Size
```{r}
# If you want to be 95% confident of estimating popl mean to within a 
#sampling error of +-2.2 and sdev assumed to be 17, what is sample size req?
e = 2.2
sigma = 17
alpha = 1- 0.95
CV = qnorm(alpha/2)
paste("CV", CV)
n = (CV * sigma / e)^2
paste('Sample size req n is: ',n)
```
What advertising medium gives a brand the most credibility in influencing brand decisions? According to survey 37% of millenials point to TV.
```{r}
phat = 0.37
# To conduct a study that would provide 95% Conf 
#tht point estimate is correct to within +- 0.05 of population prop,
# how large a sample size is req
e = 0.05
alpha = 1-0.95
CV = qnorm(alpha/2)
CV
n = (CV/e)^2 * phat*(1 - phat)

paste('Sample size req n is: ',n)

# To conduct followup study that would provide CI of 95%, point estimate correct within +-0.02 of population proportion, how large sample size is req.
e = 0.02
n = (CV/e)^2 * phat*(1 - phat)
paste('Sample size req n is: ',n)
```

```{r}
# Author examined 70 reports from 1995 to 1996. They classified 30 of reports as fav to drugs, 17 neutral and 23 critical.

# Construct 95% CI for true prop of all fav reports
phat = 30/70
n = 70
phat*n > 10
(1-phat)*n > 10
n > max(10/phat, 10/(1-phat))
# n is large enough
alpha = 1 - 0.95
CV = qnorm(alpha/2)
paste("CV", CV)
SE = sqrt(phat*(1-phat)/n)
paste("SE", SE)
LL = phat - CV * SE
UL = phat + CV * SE
paste("LL", UL)
paste("UL", LL)

## Construct 99% CI for true prop of all neutral reports
phat = 17/70
n = 70
phat*n > 10
(1-phat)*n > 10
n > max(10/phat, 10/(1-phat))
# n is large enough
alpha = 1 - 0.99
CV = qnorm(alpha/2)
paste("CV", CV)
SE = sqrt(phat*(1-phat)/n)
paste("SE", SE)
LL = phat - CV * SE
UL = phat + CV * SE
paste("LL", UL)
paste("UL", LL)
```
If bottled water distr wants to estimate within 95% C, mean amt of water in 1 gallon bottle to within +-0.003 gallons. Assume SD = 0.03gallons, what is sample size needed
```{r}
e = 0.003 
sigma = 0.03
alpha = 1 - 0.95
CV = qnorm(alpha/2)
paste("CV", CV)

n = (CV * sigma / e)^2
paste('Sample size req n is: ',n)
```
A corporate fleet which operates and manages car rentals for company employees
found that the tire lifetime for their vehicles is normally distributed with a mean of 50,000 miles and standard deviation of 6500 miles.

(a) What are the expected values for the mean and standard deviation of the sampling
distribution given a sample size of 50.
```{r}
mu = 50000
sigma = 6500
n=50
Xbar = mu =50000
paste("E(Xbar):",Xbar)
s = sigma / sqrt(n)
paste("SDev: ",s)
```

(b) What is the probability that the sample mean lifetime for these 50 vehicles exceeds
52,000?
```{r}
# P(Xbar > 52000) = 1 - P(Xbar < 52000)
1 - pnorm(52000,Xbar, s)
```
What is the probability that the sample mean lifetime for those 50 vehicles is less than
45,000?
```{r}
# P(Xbar < 45000)
pnorm(45000,Xbar, s)
```
In a telephone survey of 1007 adults conducted by the History Channel and Roper Starch between October 13 and November 2, 1998, 735 respondents stated that they believe that a conspiracy was definitely or probably behind the assassination of President John F. Kennedy on November 22, 1963, in Dallas. Construct a 99% confidence interval for the true proportion of U.S. adults who believe that a conspiracy was behind the assassination.
```{r}
n = 1007
phat =  735/n
phat * n > 10
(1-phat)*n > 10
n > max(10/phat, 10/(1-phat))
# n is large enough CI is reliable

# Construct 99% CI
# Assume phat is normally distributed
alpha = 1 - 0.99
CV = qnorm(alpha/2)
paste("CV", CV)
SE = sqrt(phat*(1 - phat)/n)
paste("SE", SE)

LL = phat - CV * SE
UL = phat + CV * SE
paste("LL", UL)
paste("UL", LL)
```
A quality characteristic of interest for a tea-bag-lling process is the weight of the
tea in the individual bags. The label weight on the package indicates that the mean amount
is 5.37 grams of tea in a bag. Problems arise if the bags are under-lled or if the mean
amount of tea in a bag exceeds the label weight. The accompanying data are the weights,
in grams, of a sample of 50 tea bags produced in one hour by a single machine (data le
Teabags.csv).
(a) Construct a 95% condence interval estimate for the population mean weight of the tea
bags.
Critical Value for 95 % condence interval is
```{r}
library(e1071)
library(car)
# Read Teabags csv file
teabag = read.csv('Teabags.csv')
# Calculate the mean of sample
mean(teabag$Teabags)
# Calculate standard deviation of sample
sd(teabag$Teabags)
hist(teabag$Teabags)
qqPlot(teabag$Teabags)

Xbar = mean(teabag$Teabags)
s = sd(teabag$Teabags)
n = 50

n
10*(skewness(teabag$Teabags))^2
10*(abs(kurtosis(teabag$Teabags)))
n > max (10*(skewness(teabag$Teabags))^2, 10*(abs(kurtosis(teabag$Teabags))))
n > max(10*((skewness(teabag$Teabags))^2), (10*(abs(kurtosis(teabag$Teabags)))))
#95% CI
alpha = 1 - 0.95
# SInce sigma is unknown we go for t distribution
CV = qt(alpha/2, n-1)
paste("CV", CV)

SE = s/sqrt(n)
paste("SE", SE)

LL = Xbar - CV * SE
UL = Xbar + CV * SE
paste("LL", UL)
paste("UL", LL)


```
(d) Explain how to understand the 95% of the interval via simulation. Use simula-
tion in your answer. Realize, of course, that the distributions that produce the
data cannot possibly be normal, but do the simulation using normal distribution
anyway.
```{r}
set.seed(1234)

mu = mean(teabag$Teabags) #inch # This is a true population parameter.
sigma = sd(teabag$Teabags) #inch # This is a true population parameter.

n = length(teabag$Teabags) # sample size
nsim = 10000 # Number of samples. Each sample is size n.

ntot = n*nsim

rv = rnorm(ntot, mu, sigma) # Generating many random numbers
rvm = matrix(rv, ncol=n, byrow=TRUE) # Convert data to a nsim*n matrix

xbars = apply(rvm,1, mean) # It's like finding rowMeans(rvm)

alpha = 0.05

lower = xbars - qnorm(1-alpha/2)*sigma/sqrt(n)
upper = xbars + qnorm(1-alpha/2)*sigma/sqrt(n)

check  = (lower < mu) & (mu < upper) # Check if confidence interval contains true mu.

head(check)

head(cbind(xbars, sigma, n, lower, upper, mu, check), 3)

# What proportion of the created confidence intervals contain mu?
mean(check)
```

```{r}
set.seed(1234)

Xbar = mean(teabag$Teabags)
s = sd(teabag$Teabags)
n = 50

nsim = 1000 # Number of samples. Each sample is size n.

ntot = n*nsim

#Assuming normal distribution
rv = rnorm(ntot, Xbar, s) # Generating many random numbers
rvm = matrix(rv, ncol=n, byrow=TRUE) # Convert data to a nsim*n matrix

#1 is for row and 2 stands for column
#xbar 10000 rows returned
xbars = apply(rvm,1, mean) # It's like finding rowMeans(rvm)

#95% of the interval 
alpha = 1-0.95

lower = xbars - qnorm(alpha/2)*s/sqrt(n)
upper = xbars + qnorm(alpha/2)*s/sqrt(n)

check  = (lower < Xbar) & (Xbar < upper) # Check if confidence interval contains true mu.

head(check)


## An estimate of the true confidence level of the ordinary confidence interval:
#cbind is like combine can use dataframe also
#each row is sample
head(cbind(xbars, s, n, lower, upper, xbar, check), 20) 
# What proportion of the created confidence intervals contain mu?
paste("Mean: ",mean(check))

hist(teabag$Teabags)
qqPlot(teabag$Teabags)
```
A survey by a national association is designed to estimate the proportion
of member organizations that have both talent and development programs in place to
drive human-capital management as well as the member organizations' mean annual
employee turnover cost. A random sample of 100 member organizations reveals the
following:
Frontline employee turnover cost: X
= $13,200, S = $1200. Forty-four member organizations have
both talent and development programs in place to drive human-capital
management.
Construct a 99% condence interval estimate for the population mean front-line em-
ployee turnover cost of member organizations.
```{r}
Xbar = 13200
s = 1200
n = 100

#99% CI
alpha = 1 - 0.99
# SInce sigma is unknown we go for t distribution
CV = qt(alpha/2, n-1)
paste("CV", CV)

SE = s/sqrt(n)
paste("SE", SE)

LL = Xbar - CV * SE
UL = Xbar + CV * SE
paste("LL", UL)
paste("UL", LL)
```

Construct a 99% condence interval estimate for the population proportion of member
organizations that have both talent and development programs in place to drive human-
capital management.
```{r}
# 99% condence interval estimate
n = 100
phat = 44/n
phat * n > 10
(1-phat)*n > 10
n > max(10/phat, 10/(1-phat))
# n is large enough CI is reliable

# Construct 99% CI
# Assume phat is normally distributed
alpha = 1 - 0.99
CV = qnorm(alpha/2)
paste("CV", CV)
SE = sqrt(phat*(1 - phat)/n)
paste("SE", SE)

LL = phat - CV * SE
UL = phat + CV * SE
paste("LL", UL)
paste("UL", LL)
```
(c) What sample size is needed to have 90% condence of estimating the population mean
front-line employee turnover cost to within plus or $300?
```{r}
e = 300
# 90% CI
alpha = 1- 0.90
CV = qnorm(alpha/2)
paste("CV", CV)
# sample size
sn = (CV * s / e)^2
paste('sample size is needed: ',sn)
```

(d) How many member organizations need to be selected to have 95% condence of estimat-
ing the population proportion of organizations that have both talent and development
programs in place to drive human-capital management to within plus or $0:035? As-
sume that the association's survey provides a reliable point estimate of the proportion.
Sampling error e = $0:035
```{r}
e = 0.035
# 95% CI
alpha = 1-0.95
CV = qnorm(alpha/2)
paste("CV", CV)
# sample size
sn = (CV/e)^2 * phat * (1-phat)
paste('sample size is needed: ',sn)
```

