---
title: "What Confidence Interval Means?"
author: "Alireza Sheikh-Zadeh"
date: "September 27, 2018"
output:
  html_document: default
  word_document: default
---

A **confidence level** specifies the probability that our particular sample's interval estimate will in fact contain the population parameter.


**Common Misinterpretations of Confidence Level**

Many statisticians are at pains to remind non-statisticians that confidence levels are very easily misunderstood. One common misinterpretation is that your interval estimate will hold a certain percentage of all cases in the population. For example, one could mistakenly conclude that 95% of all of the weights of adult males in the United States will be between 150 and 165 pounds. A correct interpretation would be: We can be 95% confident that our interval estimate of 150 to 165 pounds will in fact contain the true population value. 

Simulation example: Let's simulate normally distributed random samples for the height of adult women in the US. We generate a population data by simulation. Then we find 95% confidence interval for multiple samples taken from population. The sample size is n.  


```{r}

set.seed(1234)

# Heights are normally distributed and the true parameters are as follow: 
mu = 65 #inch # This is a true population parameter.
sigma = 3.5 #inch # This is a true population parameter.

n = 30 # sample size
nsim = 100000 # Number of samples. Each sample is size n.
ntot = n*nsim

rv = rnorm(ntot, mu, sigma) # Generating many random numbers
rvm = matrix(rv, ncol=n, byrow=TRUE) # Convert data to a nsim*n matrix

xbars = apply(rvm,1, mean) # It's like finding rowMeans(rvm)

alpha = 0.05

lower = xbars - qnorm(1-alpha/2)*sigma/sqrt(n)
upper = xbars + qnorm(1-alpha/2)*sigma/sqrt(n)

check  = (lower < mu) & (mu < upper) # Check if confidence interval contains true mu.

head(check)


## An estimate of the true confidence level of the ordinary confidence interval:
head(cbind(xbars, sigma, n, lower, upper, mu, check), 20) 

# What proportion of the created confidence intervals contain mu?
mean(check)

```


**Understanding the confidence interval for mu: Z versus t**

Let's assume sigma is unknown. 


```{r}
# We need to find sample standard deviations. 
s = apply(rvm,1, sd)

# First use z critical value

lower = xbars - qnorm(1-alpha/2)*s/sqrt(n)
upper = xbars + qnorm(1-alpha/2)*s/sqrt(n)

check  = (lower < mu) & (mu < upper) # Check if confidence interval contains true mu.
mean(check)

# Second use t critical value

lower = xbars - qt(1-alpha/2, n-1)*s/sqrt(n)
upper = xbars + qt(1-alpha/2, n-1)*s/sqrt(n)

check  = (lower < mu) & (mu < upper) # Check if confidence interval contains true mu.
mean(check)
## Notice that the coverage rate is closer to 95% but not perfect.

head(cbind(xbars, s, n, lower, upper, mu, check), 10) 

```

```{r}
# Example 3 in the Note

a <- c(1, 4, 4, 4, 5, 5, 5, 8)

b <- 1:8

# Part a:
xbar = mean(a)
xbar
n = length(a)
s = sd(a)

se = s/sqrt(n)

alpha = 0.05

lower = xbar - qt(1-alpha/2, n-1)*se 
upper = xbar + qt(1-alpha/2, n-1)*se 
lower
upper
```

```{r}
b <- 1:8

# Part b:
xbar = mean(b)
xbar
n = length(b)
s = sd(b)

se = s/sqrt(n)

alpha = 0.05

lower = xbar - qt(1-alpha/2, n-1)*se 
upper = xbar + qt(1-alpha/2, n-1)*se 
lower
upper
```


